---
title: "Jun Pan DATA 607 wk5 assignment final.rmd"
author: "Jun Pan"
date: "September 28, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

write csv file
```{r}
library(dplyr)
library(readr)
library(tidyr)
library(knitr)
library(data.table)
```
```{r}
flight = rbind(c("Airline","Status","Los Angeles", "Phoenix", "San Diego", "San Francisco", "Seattle"),
                c("ALASKA", "ontime", "497","221","212","503","1841"),
                c("ALASKA", "delayed", "62","12","20","102","305"),
                c("AM WEST", "ontime", "694", "4840","383","320","201"),
                c("AM WEST", "delayed", "117","415","65","129","61"))
flight = data.frame(flight)
```


```{r}
flight
```

```{r}
write.csv(flight, file = "flight.csv")
```

```{r}
flight <- read.csv(file="C:/Users/tbao/Desktop/CUNY MSDS notes/607/Week 5/week 5 assignment/flight.csv")
```

```{r}
flight
```
tidy data
```{r}
library(tidyr)
```

```{r}
airlines_transform1 <- tidyr::gather(flight, "city", "count", 3:7)
```


display
```{r}
kable(airlines_transform1, format="markdown")
```


```{r}
airlines_transform2 <- tidyr::spread(airlines_transform1, Status, Count)
```

```{r}
kable(airlines_transform2, format="markdown")
```
The following data show the maxium, minum, average, standard deviation, median and quatile of delayed flight 
```{r}
delay_summary <- airlines_transform2 %>% dplyr::group_by(Airline) %>% 
  dplyr::summarise(max=max(delayed), 
                   min=min(delayed),
                   avg=mean(delayed),
                   sd_delay=sd(delayed),
                   median=median(delayed),
                   IQR=IQR(delayed))
#display
kable(delay_summary, format="markdown")
```

Create separate data.frames for Alaska Airlines and AM_West Airlines.
```{r}
Alaska <- airlines_transform2 %>% filter(Airline == 'ALASKA')
AMWest <- airlines_transform2 %>% filter(Airline == 'AMWEST')
```

```{r}
Alaska %>%
  arrange(desc(delayed))
```
The above table shows that Seattle has the most delayed fight counts in Alaska airline.


```{r}
AMWest %>%
  arrange(desc(delayed))
```
The above table shows the descending order by counts for delayed flights, we can see phoenix has the most delayed fight by AMWEST

```{r}
library(ggplot2)
```

AM West delayed counts by city
```{r}
AMWestDelays<-ggplot(data=AMWest, aes(x=City, y=delayed)) +
  geom_bar(stat="identity") + 
  geom_bar(stat="identity", color="blue", fill="white")
AMWestDelays
```


Alasak airline delayed fight counts by cities
```{r}
AlaskaDelays<-ggplot(data=Alaska, aes(x=City,y=delayed)) +
  geom_bar(stat="identity") + 
  geom_bar(stat="identity", color="Red", fill="white")
AlaskaDelays
```


